---
title: "Westchester and surrounding counties"
site: distill::distill_website
preview: img/wch-map.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(plotly)
library(sf)
library(DT)
library(here)
library(slider)
library(scales)
library(lubridate)
library(leaflet)
library(leaflet.extras)
library(gghighlight)
library(extrafont)
library(xfun)

source(here("R/utils.R"))
```

```{r xaringan-panelset, echo=FALSE}
xaringanExtra::use_panelset()
```

```{r read_in_data}
county_bound <- read_rds(here("data/county-boundary.rds"))

nys_cases <- read_csv(here("data/by-county-cases-tests-nys.csv")) %>% 
  group_by(county) %>% 
  mutate(
    across(c(new_cases, total_cases, new_tests, total_tests), 
    list(
      "avg_7" = ~ slide_dbl(.x, mean, .before = 6, .complete = TRUE),
      "avg_14" = ~ slide_dbl(.x, mean, .before = 13, .complete = TRUE)
      )
    ),
    pos_rate_avg_7 = new_cases_avg_7 / new_tests_avg_7,
    pos_rate_avg_14 = new_cases_avg_14 / new_tests_avg_14
    ) %>% 
  ungroup()

nyt_cases <- read_csv(here("data/by-county-cases-deaths-nyt.csv")) %>% 
  group_by(county, state) %>% 
  mutate(
    across(where(is.numeric), 
    list(
      "avg_7" = ~ slide_dbl(.x, mean, .before = 6, .complete = TRUE),
      "avg_14" = ~ slide_dbl(.x, mean, .before = 13, .complete = TRUE)
      )
    )) %>% 
  ungroup() %>% 
  inner_join(select(county_bound, -fips), by = c("county", "state")) %>% 
  mutate(
    new_cases_per_cap = new_cases_avg_7 / total_pop * 1e5,
    tot_cases_per_cap = total_cases / total_pop,
    new_deaths_per_cap = new_deaths_avg_7 / total_pop * 1e5
    ) %>% 
  select(-geometry)
```

```{r}
recent <- nys_cases %>% 
  filter(county == "Westchester", date == max(date))

recent_nyt <- nyt_cases %>% 
  filter(county == "Westchester", date == max(date))
```

On `r month_day_year(recent$date, abbr = TRUE)`, `r comma(recent$new_cases, 1)` people tested positive for SARS-CoV2 out of `r comma(recent$new_tests, 1)` tests reported in Westchester. This is a test positivity rate of `r percent(recent$pos_rate, 0.1)`. `r n2w(recent_nyt$new_deaths, and = TRUE, hyphen = FALSE, cap = TRUE)` new deaths due to Covid-19 were reported on `r month_day_year(recent_nyt$date, abbr = TRUE)`. Since the beginning of the pandemic, `r comma(recent$total_cases, 1)` people have tested positive and `r comma(recent_nyt$total_deaths, 1)` have died in Westchester County.

<br>

<div class = "l-body-outset">
::::: {.panelset}
::: {.panel}
[Cases]{.panel-name}

```{r county_case_bar, layout="l-body-outset"}
to_plot <- nys_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, new_cases)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "New cases: ", comma(new_cases, 1), "<br>",
        "7-day average: ", comma(new_cases_avg_7, 1)
        ),  "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = new_cases_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "1 month",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_number_si()) +
  labs(
    title = "New cases reported in Weschester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(hovermode = "x unified")
```
:::

::: {.panel}
[Tests]{.panel-name}

```{r county_test_bar, layout="l-body-outset"}
to_plot <- nys_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, new_tests)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "New tests: ", comma(new_tests, 1), "<br>",
        "7-day average: ", comma(new_tests_avg_7, 1)
        ),  "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = new_tests_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "1 month",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_number_si()) +
  labs(
    title = "New tests reported in Weschester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(hovermode = "x unified")
```
:::

::: {.panel}
[Deaths]{.panel-name}

```{r county_death_bar, layout="l-body-outset"}
to_plot <- nyt_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, new_deaths)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "New deaths: ", comma(new_deaths, 1), "<br>",
        "7-day average: ", comma(new_deaths_avg_7, 1)
        ),  "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = new_deaths_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "1 month",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_number_si()) +
  labs(
    title = "New deaths reported in Weschester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(hovermode = "x unified")
```
:::

::: {.panel}
[Positivity Rate]{.panel-name}

```{r county_pos_rate_bar, layout="l-body-outset"}
to_plot <- nys_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, pos_rate)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "Positivity rate: ", percent(pos_rate, 0.1), "<br>",
        "7-day average: ", percent(pos_rate_avg_7, 0.1)
        ),  "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = pos_rate_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "1 month",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_percent(1), limits = c(0, 0.507)) +
  labs(
    title = "Test postivity rate in Weschester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(hovermode = "x unified")
```
:::
:::::
</div>

```{r westchester-table, layout="l-body-outset"}
deaths <- nyt_cases %>% 
  filter(county == "Westchester") %>% 
  select(date, new_deaths, total_deaths)

nys_cases %>% 
  filter(county == "Westchester") %>% 
  select(date, new_cases, new_tests, total_cases, pos_rate) %>% 
  left_join(deaths, by = "date") %>% 
  arrange(desc(date)) %>% 
  mutate(
    across(where(is.numeric), ~ if_else(is.na(.x), 0, .x)),
    date = month_day_year(date, abbr = TRUE)
    ) %>% 
  select(
    `Report date` = date,
    `New cases` = new_cases,
    `New tests` = new_tests,
    `New deaths` = new_deaths,
    `Test positivity rate` = pos_rate,
    `Total cases` = total_cases,
    `Total deaths` = total_deaths
    ) %>% 
  datatable(
    class = "compact hover",
    extensions = "Scroller",
    options = list(
      dom = "ti",
      deferRender = TRUE,
      scrollY = 250,
      scroller = TRUE,
      columnDefs = list(
        list(orderData = 0, targets = 1),
        list(visible = FALSE, targets = 0),
        list(className = "dt-head-left", targets = 1)
        )
      )
    ) %>% 
  formatPercentage(5, digits = 1) %>% 
  formatRound(c(2:4, 6:7), digits = 0) %>% 
  formatStyle(1, fontFamily = "IBM Plex Sans") %>% 
  formatStyle(2:7, fontFamily = "Roboto Mono")
```

## Metro area

```{r}
recent_metro_nyt <- nyt_cases %>% 
  filter(date == max(date), county != "Westchester") %>% 
  arrange(desc(new_cases_per_cap))

tot_metro_nyt <- nyt_cases %>% 
  filter(date == max(date), county != "Westchester") %>% 
  arrange(desc(tot_cases_per_cap))

tot_metro_least <- nyt_cases %>% 
  filter(date == max(date), county != "Westchester") %>% 
  arrange(tot_cases_per_cap)
```

In the New York metropolitan area, the county with the most new cases per capita in the past seven days is **`r recent_metro_nyt$county[1]` County**, where there have been an average of `r comma(recent_metro_nyt$new_cases_avg_7[1], 1)` cases per day reported in the past week (`r comma(recent_metro_nyt$new_cases_per_cap[1], 1)` per 100,000 residents). **`r recent_metro_nyt$county[2]` County** is another recent hot spot, with `r comma(recent_metro_nyt$new_cases_avg_7[2], 1)` average daily cases cases reported in the past week (`r comma(recent_metro_nyt$new_cases_per_cap[2], 1)` per 100,000). In **Westchester County**, `r comma(recent_nyt$new_cases_per_cap[1], 1)` cases per 100,000 residents have been reported in the past week (`r comma(recent_nyt$new_cases_avg_7[1], 1)` new cases per day).

Throughout the pandemic, the counties that have had the largest share of their population test positive for coronavirus are **`r tot_metro_nyt$county[1]`** (`r pretty_frac(tot_metro_nyt$tot_cases_per_cap[1])`) and **`r tot_metro_nyt$county[2]`** (`r pretty_frac(tot_metro_nyt$tot_cases_per_cap[2])`). The counties that have had the fewest cases per capita are **`r tot_metro_least$county[1]`** (`r pretty_frac(tot_metro_least$tot_cases_per_cap[1])`) and **`r tot_metro_least$county[2]`** (`r pretty_frac(tot_metro_least$tot_cases_per_cap[2])`). In **Westchester County**, `r pretty_frac(recent_nyt$tot_cases_per_cap)` residents have tested positive since the beginning of the pandemic (`r comma(recent_nyt$total_cases, 1)`).

### New cases per capita

```{r metro_cases_map, out.height = "600px"}
to_map <- county_bound %>% 
  inner_join(nyt_cases, by = c("county", "state")) %>% 
  filter(date == max(date)) %>% 
  mutate(
    label = paste0(
      "<b>", county, "</b><br>",
      "Avg. daily cases: ", comma(new_cases_avg_7, 1), "<br>",
      "Per 100k: ", number(new_cases_per_cap, 0.1)
      )
    )

cd_pal <- colorNumeric(
  palette = "YlOrRd",
  domain = 0:100
  )

leaflet(to_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    weight = 1.5,
    color = "white",
    fillColor = ~cd_pal(new_cases_per_cap),
    fillOpacity = 0.8,
    label = lapply(to_map$label, htmltools::HTML),
    highlight = highlightOptions(
      weight = 2,
      color = "#666666",
      opacity = 0.7,
      fillOpacity = 0.7,
      bringToFront = TRUE
      )
    ) %>%
  addLegend(
    pal = cd_pal,
    values = to_map$new_cases_per_cap,
    opacity = 0.7,
    bins = c(20, 40, 60, 80, 100),
    position = "bottomright",
    title = "Avg. daily cases per 100k",
  ) %>% 
  setView(lng = -73.75, lat = 41.3, zoom = 8) %>% 
  addResetMapButton()
```

<br>

```{r county-table, layout="l-body-outset", out.height="510px"}
to_map %>% 
  st_drop_geometry() %>% 
  arrange(desc(new_cases_per_cap)) %>% 
  mutate(tot_cases_per_cap = pretty_frac(tot_cases_per_cap)) %>% 
  select(
    `County` = county,
    `Avg daily cases per 100k` = new_cases_per_cap,
    `Avg daily cases` = new_cases_avg_7,
    `Total cases` = total_cases,
    `Share of pop` = tot_cases_per_cap,
    `Total pop` = total_pop.x
    ) %>% 
  datatable(
    class = "compact hover",
    extensions = "Scroller",
    options = list(
      dom = "fti",
      deferRender = TRUE,
      scrollY = 250,
      scroller = TRUE,
      columnDefs = list(
        list(visible = FALSE, targets = 0),
        list(className = "dt-head-left", targets = 1),
        list(className = "dt-center", targets = 5)
        )
      )
    ) %>% 
  formatRound(2, digits = 1) %>% 
  formatRound(c(3:4, 6), digits = 0) %>% 
  formatStyle(1, fontFamily = "IBM Plex Sans") %>% 
  formatStyle(2:6, fontFamily = "Roboto Mono")
```

<br>

```{r small_mult_cases_per_cap, layout="l-page", dpi=300, dev="svg", fig.height=6, fig.width=12}
nyt_cases %>% 
  mutate(
    county = fct_reorder(county, -new_cases_per_cap, mean, na.rm = TRUE),
    county = fct_relevel(county, "Westchester")
    ) %>%
  ggplot(aes(date, new_cases_per_cap, color = county)) +
  geom_line(color = "steelblue", size = 1) +
  gghighlight(
    use_direct_label = FALSE,
    unhighlighted_params = list(size = 0.2)
    ) +
  facet_wrap(vars(county), nrow = 3) +
  labs(
    title = "New cases per 100,000 residents",
    subtitle = str_wrap("Trend lines for all counties in the metro area are drawn in gray to compare with trend of each county drawn in blue", 65),
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_family = "IBM Plex Sans") +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 14, hjust = 0),
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 12)
    )
```


