---
title: "Westchester and surrounding counties"
site: distill::distill_website
preview: img/wch-map.png
output:
  distill::distill_article:
    toc: true
    toc_depth: 2
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
xaringanExtra::use_panelset()
```

```{r}
recent <- nys_cases %>% 
  filter(county == "Westchester", date == max(date))

recent_nyt <- nyt_cases %>% 
  filter(county == "Westchester", date == max(date))

cases_per_cap <- pretty_frac(recent$total_cases / recent_nyt$total_pop)
death_per_cap <- pretty_frac(recent_nyt$total_deaths / recent_nyt$total_pop)
```

<span class="recent-update">Most recent data: `r month_day_year(recent$date, abbr = TRUE)`</a>

On `r month_day_year(recent$date, abbr = TRUE)`, `r comma(recent$new_cases, 1)` people tested positive for the novel coronavirus out of `r comma(recent$new_tests, 1)` tests reported in Westchester. This is a test positivity rate of `r percent(recent$pos_rate, 0.1)`. `r n2w(recent_nyt$new_deaths, and = TRUE, hyphen = FALSE, cap = TRUE)` new deaths due to Covid-19 were reported on `r month_day_year(recent_nyt$date, abbr = TRUE)`. Since the beginning of the pandemic, `r comma(recent$total_cases, 1)` people have tested positive (`r cases_per_cap`) and `r comma(recent_nyt$total_deaths, 1)` have died in Westchester County (`r death_per_cap`).

<br>

<div class = "l-body-outset">
::::: {.panelset}
::: {.panel}
[Cases]{.panel-name}

```{r county_case_bar, layout="l-body-outset"}
to_plot <- nys_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, new_cases)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "New cases: ", comma(new_cases, 1), "<br>",
        "7-day average: ", comma(new_cases_avg_7, 1)
        ),  "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = new_cases_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "2 months",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_number_si()) +
  labs(
    title = "New cases reported in Westchester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(
    hovermode = "x unified",
    annotations = list(
      x = 1, y = -0.175,
      text = "Source: <a href='https://westchester-covid.mattherman.info/about.html#new-york-state-statewide-covid-19-testing' target='_self'>NYS Dept of Health</a>", 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
:::

::: {.panel}
[Tests]{.panel-name}

```{r county_test_bar, layout="l-body-outset"}
to_plot <- nys_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, new_tests)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "New tests: ", comma(new_tests, 1), "<br>",
        "7-day average: ", comma(new_tests_avg_7, 1)
        ),  "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = new_tests_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "2 months",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_number_si()) +
  labs(
    title = "New tests reported in Westchester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(
    hovermode = "x unified",
    annotations = list(
      x = 1, y = -0.175,
      text = "Source: <a href='https://westchester-covid.mattherman.info/about.html#new-york-state-statewide-covid-19-testing' target='_self'>NYS Dept of Health</a>", 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
:::

::: {.panel}
[Positivity Rate]{.panel-name}

```{r county_pos_rate_bar, layout="l-body-outset"}
to_plot <- nys_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, pos_rate)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "Positivity rate: ", percent(pos_rate, 0.1), "<br>",
        "7-day average: ", percent(pos_rate_avg_7, 0.1)
        ),  "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = pos_rate_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "2 months",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_percent(1), limits = c(0, 0.507)) +
  labs(
    title = "Test postivity rate in Westchester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(
    hovermode = "x unified",
    annotations = list(
      x = 1, y = -0.175,
      text = "Source: <a href='https://westchester-covid.mattherman.info/about.html#new-york-state-statewide-covid-19-testing' target='_self'>NYS Dept of Health</a>", 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
:::

::: {.panel}
[Deaths]{.panel-name}

```{r county_death_bar, layout="l-body-outset"}
to_plot <- nyt_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, new_deaths)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "New deaths: ", comma(new_deaths, 1), "<br>",
        "7-day average: ", comma(new_deaths_avg_7, 1)
        ),  "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = new_deaths_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "2 months",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_number_si()) +
  labs(
    title = "New deaths reported in Westchester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(
    hovermode = "x unified",
    annotations = list(
      x = 1, y = -0.175,
      text = "Source: <a href='https://westchester-covid.mattherman.info/about.html#coronavirus-covid-19-data-in-the-united-states' target='_self'>NY Times</a>", 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
:::

:::::
</div>

```{r westchester-table, layout="l-body-outset"}
deaths <- nyt_cases %>% 
  filter(county == "Westchester") %>% 
  select(date, new_deaths, total_deaths)

nys_cases %>% 
  filter(county == "Westchester") %>% 
  select(date, new_cases, new_tests, total_cases, pos_rate) %>% 
  left_join(deaths, by = "date") %>% 
  arrange(desc(date)) %>% 
  mutate(
    across(where(is.numeric), ~ if_else(is.na(.x), 0, .x)),
    date = month_day_year(date, abbr = TRUE)
    ) %>% 
  select(
    `Report date` = date,
    `New cases` = new_cases,
    `New tests` = new_tests,
    `New deaths` = new_deaths,
    `Test positivity rate` = pos_rate,
    `Total cases` = total_cases,
    `Total deaths` = total_deaths
    ) %>% 
  datatable(
    class = "compact hover",
    extensions = c("Scroller", "Responsive"),
    caption = tags$caption(
      withTags(div(HTML('Source: <a href="https://westchester-covid.mattherman.info/about.html#new-york-state-statewide-covid-19-testing">NYS Dept of Health and NY Times</a>')))
    ),
    options = list(
      dom = "t",
      deferRender = TRUE,
      scrollY = 250,
      scroller = TRUE,
      columnDefs = list(
        list(orderData = 0, targets = 1),
        list(visible = FALSE, targets = 0),
        list(className = "dt-head-left", targets = 1)
        )
      )
    ) %>% 
  formatPercentage(5, digits = 1) %>% 
  formatRound(c(2:4, 6:7), digits = 0) %>% 
  formatStyle(1, fontFamily = "IBM Plex Sans") %>% 
  formatStyle(2:7, fontFamily = "Roboto Mono")
```


## Metro area cases rates

```{r}
recent_metro_nyt <- nyt_cases %>%
  filter(date == max(date), county != "Westchester") %>% 
  arrange(desc(new_cases_per_cap))

tot_metro_nyt <- nyt_cases %>% 
  filter(date == max(date), county != "Westchester") %>% 
  arrange(desc(tot_cases_per_cap))

tot_metro_least <- nyt_cases %>% 
  filter(date == max(date), county != "Westchester") %>% 
  arrange(tot_cases_per_cap)
```

In the New York metropolitan area, the county with the most new cases per capita in the past seven days is **`r recent_metro_nyt$county[1]` County**, where there have been an average of `r comma(recent_metro_nyt$new_cases_avg_7[1], 1)` cases per day reported in the past week (`r comma(recent_metro_nyt$new_cases_per_cap[1], 1)` per 100,000 residents). **`r recent_metro_nyt$county[2]` County** is another recent hot spot, with `r comma(recent_metro_nyt$new_cases_avg_7[2], 1)` average daily cases cases reported in the past week (`r comma(recent_metro_nyt$new_cases_per_cap[2], 1)` per 100,000). In **Westchester County**, `r comma(recent_nyt$new_cases_per_cap[1], 1)` cases per 100,000 residents have been reported in the past week (`r comma(recent_nyt$new_cases_avg_7[1], 1)` new cases per day).

Throughout the pandemic, the counties that have had the largest share of their population test positive for coronavirus are **`r tot_metro_nyt$county[1]`** (`r pretty_frac(tot_metro_nyt$tot_cases_per_cap[1])`) and **`r tot_metro_nyt$county[2]`** (`r pretty_frac(tot_metro_nyt$tot_cases_per_cap[2])`). The counties that have had the fewest cases per capita are **`r tot_metro_least$county[1]`** (`r pretty_frac(tot_metro_least$tot_cases_per_cap[1])`) and **`r tot_metro_least$county[2]`** (`r pretty_frac(tot_metro_least$tot_cases_per_cap[2])`). In **Westchester County**, `r pretty_frac(recent_nyt$tot_cases_per_cap)` residents have tested positive since the beginning of the pandemic.

### New cases per capita

```{r metro_cases_map, out.height = "600px"}
to_map <- county_bound %>% 
  inner_join(nyt_cases, by = c("county", "state")) %>% 
  filter(date == max(date)) %>% 
  mutate(
    label = paste0(
      "<b>", county, "</b><br>",
      "Avg. daily cases: ", comma(new_cases_avg_7, 1), "<br>",
      "Per 100k: ", number(new_cases_per_cap, 0.1)
      )
    )

county_pal <- colorNumeric(
  palette = "YlOrRd",
  domain = 0:100
  )

leaflet(to_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    group = "Cases per 100k",
    weight = 1.5,
    color = "white",
    fillColor = ~county_pal(new_cases_per_cap),
    fillOpacity = 0.8,
    label = lapply(to_map$label, htmltools::HTML),
    highlight = highlightOptions(
      weight = 2,
      color = "#666666",
      opacity = 0.7,
      fillOpacity = 0.7,
      bringToFront = TRUE
      )
    ) %>%
  addLegend(
    pal = county_pal,
    values = 0:100,
    opacity = 0.7,
    bins = c(0, 25, 50, 75, 100),
    position = "bottomright",
    title = "Avg. daily cases per 100k"
    ) %>% 
  setView(lng = -73.75, lat = 41.3, zoom = 8) %>% 
  addResetMapButton()
```

<br>

```{r county-table, layout="l-body-outset", out.height="510px"}
to_map %>% 
  st_drop_geometry() %>% 
  arrange(desc(new_cases_per_cap)) %>% 
  mutate(tot_cases_per_cap = pretty_frac(tot_cases_per_cap)) %>% 
  select(
    `County` = county,
    `Avg daily cases per 100k` = new_cases_per_cap,
    `Avg daily cases` = new_cases_avg_7,
    `Total cases` = total_cases,
    `Share of pop` = tot_cases_per_cap,
    `Total pop` = total_pop.x
    ) %>% 
  datatable(
    class = "compact hover",
    extensions = c("Scroller", "Responsive"),
    caption = tags$caption(
      style = 'caption-side: bottom; text-align: right;',
      withTags(div(HTML('Source: <a href="https://westchester-covid.mattherman.info/about.html#coronavirus-covid-19-data-in-the-united-states">NY Times</a>')))
    ),
    options = list(
      dom = "ft",
      deferRender = TRUE,
      scrollY = 250,
      scroller = TRUE,
      columnDefs = list(
        list(visible = FALSE, targets = 0),
        list(className = "dt-head-left", targets = 1),
        list(className = "dt-center", targets = 5)
        )
      )
    ) %>% 
  formatRound(2, digits = 1) %>% 
  formatRound(c(3:4, 6), digits = 0) %>% 
  formatStyle(1, fontFamily = "IBM Plex Sans") %>% 
  formatStyle(2:6, fontFamily = "Roboto Mono")
```

The timing of outbreaks in counties in the New York metro area has varied. Some counties such as Rockland, Westchester, and Nassau had severe first waves in spring 2020. By contrast, other counties such as New Haven, Fairfield, Litchfield, had relatively mild first waves in the spring and more severe second waves in fall 2020. The chart below shows the number of cases per 100,000 residents from March 2020 through the present in each county.

```{r small_mult_cases_per_cap, layout="l-page", dpi=300, dev="svg", fig.height=6, fig.width=12}
nyt_cases %>% 
  mutate(
    county = fct_reorder(county, -new_cases_per_cap, mean, na.rm = TRUE),
    county = fct_relevel(county, "Westchester")
    ) %>%
  ggplot(aes(date, new_cases_per_cap, color = county)) +
  geom_line(color = "steelblue", size = 0.8) +
  gghighlight(
    use_direct_label = FALSE,
    unhighlighted_params = list(size = 0.2)
    ) +
  scale_x_date(expand = expansion(add = c(0, 2))) +
  facet_wrap(vars(county), nrow = 3, strip.position = "bottom") +
  labs(
    title = "New cases per 100,000 residents",
    subtitle = str_wrap("Trend lines for all counties in the metro area are drawn in gray to compare with trend of each county drawn in blue", 65),
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_family = "IBM Plex Sans") +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank(),
    strip.text = element_text(size = 12, hjust = 0.1, color = "gray15", family = "Roboto Condensed"),
    axis.text.x = element_text(size = 8, color = "gray45"),
    strip.placement = "outside",
    plot.title = element_text(face = "bold", size = 18),
    plot.subtitle = element_text(size = 12),
    panel.spacing.y = unit(12, "points")
    #panel.border = element_rect(color = "gray95", fill = NA, size = 0.5)
    )
```

<br>

## New York State micro-clusters

```{r}
wch_cluster <- hotspot %>% 
  st_transform(2263) %>% 
  st_filter(mun_bound)

red <- wch_cluster %>% 
  filter(zone == "red") %>% 
  nrow()

orange <- wch_cluster %>% 
  filter(zone == "Orange") %>% 
  nrow()

yellow <- wch_cluster %>% 
  filter(zone == "Yellow") %>% 
  nrow()

red_text <- case_when(
  red == 0 ~ "are no red zones",
  red == 1 ~ "is one red zone",
  TRUE ~ paste("are", n2w(red), "red zones")
  )

orange_text <- case_when(
  orange == 0 ~ "no orange zones",
  orange == 1 ~ "one orange zone",
  TRUE ~ paste(n2w(orange), "orange zones")
  )

yellow_text <- case_when(
  yellow == 0 ~ "no yellow zones",
  yellow == 1 ~ "one yellow zone",
  TRUE ~ paste(n2w(yellow), "yellow zones")
  )
```

<span class="recent-update">Most recent data: `r month_day_year(max(wch_cluster$date), abbr = TRUE)`</span>

[Governor Cuomo announced a "micro-cluster strategy"](https://www.wgrz.com/article/news/health/coronavirus/cluster-action-initiative-will-become-more-specific/71-cf1e7e41-850c-48c0-8887-82c610258d2a) in October 2020, in which small geographic areas are targeted with [enhanced restrictions](https://forward.ny.gov/cluster-action-initiative) on gatherings, indoor dining, and other activities. Yellow, orange, and red zones have different restriction and the clusters are determined using a combination of test positivity rate, hospital admissions, and hospital capacity.

Currently, there `r red_text`, `r orange_text`, and `r yellow_text` in Westchester County.

```{r hotspot-map, out.height = "600px"}
hotspot_to_map <- hotspot %>% 
  mutate(
    label = paste0(
      "<b>", cluster, "</b><br>",
      "7-day avg positivity rate: ",  percent(pos_rate, accuracy = 0.1, scale = 1), "<br>",
      "Zone effective date: ", month_day_year(date_eff, abbr = TRUE))
    )

hotspot_pal <- colorFactor(
  palette = c("yellow2", "orange", "firebrick"),
  levels = c("Yellow", "Orange", "Red")
  )

leaflet(hotspot_to_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    data = hotspot_to_map,
    group = "NYS micro-clusters",
    weight = 1.5,
    color = "white",
    fillColor = ~hotspot_pal(zone),
    fillOpacity = 0.8,
    label = lapply(hotspot_to_map$label, htmltools::HTML),
    highlight = highlightOptions(
      weight = 2,
      color = "#666666",
      opacity = 0.7,
      fillOpacity = 0.7,
      bringToFront = TRUE
      )
    ) %>% 
  setView(lat = 41.1, lng = -73.75, zoom = 10) %>% 
  addResetMapButton()
```

<br>

## Deaths in nursing homes

```{r}
wch_nh_sum <- nh_clean %>% 
  st_drop_geometry() %>% 
  group_by(date) %>% 
  summarize(n = n(), deaths = sum(deaths_total) + 2, .groups = "drop")  # victoria house has deaths, but closed in sept

wch_deaths <- nyt_cases %>% 
  filter(date == wch_nh_sum$date, county == "Westchester")
```

<span class="recent-update">Most recent data: `r month_day_year(max(wch_nh_sum$date), abbr = TRUE)`</span>

The Covid-19 pandemic has devastated long-term care facilities. The COVID Tracking Project estimates that [38% of all people in the US who died from Covid-19](https://covidtracking.com/data/long-term-care) were residents of long-term care facilities such as nursing homes and assisted living facilities.

There have been `r comma(wch_nh_sum$deaths, 1)` confirmed or presumed deaths to due Covid-19 in nursing homes in Westchester County since the beginning of the pandemic. This is `r percent(wch_nh_sum$deaths / wch_deaths$total_deaths, 1)` of all deaths reported in Westchester.

<aside>
This is an [undercount of deaths of nursing home residents](https://www.propublica.org/article/not-mentioned-in-cuomos-coronavirus-book-how-many-nursing-home-residents-died-in-new-york) because NYS DOH only reports deaths that occurred in nursing homes. If a nursing home resident died in a hospital, it is not counted as a nursing home death.
</aside>

### Confirmed and presumed Covid-19 deaths in nursing homes
```{r nursing-home-map, out.height="600px", out.width="600px"}
to_map <- nh_clean %>%
  mutate(
    label = paste0(
      "<b>", name, "</b><br>",
      "Deaths: ", comma(deaths_total, 1), "<br>",
      "Beds: ", comma(beds, 1)
      ),
    name = fct_reorder(name, deaths_total)
    )

to_map %>%
  leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircles(
    radius = ~ sqrt(deaths_total) * 450,
    color = "steelblue",
    opacity = 0.7,
    weight = 1.5,
    fillColor = "steelblue",
    fillOpacity = 0.2,
    label = lapply(to_map$label, htmltools::HTML),
    highlightOptions = highlightOptions(sendToBack = TRUE, color = "#6fb6f2", weight = 2.5)
    ) %>%
  setView(lat = 41.1, lng = -73.75, zoom = 10) %>%
  addResetMapButton()
```

<br>

```{r nursing-home-table, layout="l-body-outset", out.height="510px"}
to_map %>% 
  st_drop_geometry() %>% 
  arrange(desc(deaths_total)) %>% 
  select(
    `Nursing Home` = name,
    City = city,
    Deaths = deaths_total,
    `Total Beds` = beds
    ) %>% 
  datatable(
    class = "compact hover",
    extensions = c("Scroller", "Responsive"),
    caption = tags$caption(
      style = 'caption-side: bottom; text-align: right;',
      withTags(div(HTML('Source: <a href="https://westchester-covid.mattherman.info/about.html#nursing-home-and-acf-covid-related-deaths-statewide">NYS DOH</a>')))
    ),
    options = list(
      dom = "ft",
      deferRender = TRUE,
      scrollY = 250,
      scroller = TRUE,
      columnDefs = list(
        list(visible = FALSE, targets = 0),
        list(className = "dt-head-left", targets = 1:2),
        list(targets = 1, render = JS(
              "function(data, type, row, meta) {",
              "return type === 'display' && data.length > 40 ?",
              "'<span title=\"' + data + '\">' + data.substr(0, 40) + '...</span>' : data;",
              "}")
             )
        )
      )
    ) %>% 
  formatRound(3:4, digits = 0) %>% 
  formatStyle(1:2, fontFamily = "IBM Plex Sans") %>% 
  formatStyle(3:4, fontFamily = "Roboto Mono")
```
