---
title: "Hospital capacity"
site: distill::distill_website
preview: img/wch-map.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
xaringanExtra::use_panelset()
```

```{r}
hosp_wch <- hosp_cap %>% 
  filter(
    year_week == max(year_week),
    county == "Westchester"
    )

hosp_wch_sum <- hosp_wch %>% 
  group_by(year_week) %>% 
  summarize(across(where(is.numeric), sum, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    bed_free = bed_capacity - bed_occupied,
    icu_free = icu_capacity - icu_occupied,
    pct_covid = covid_patients / all_patients,
    pct_bed = bed_occupied / bed_capacity,
    pct_icu = icu_occupied / icu_capacity
    )

hosp_nys <- hosp_cap %>% 
  filter(
    year_week == max(year_week),
    state == "NY"
    )

hosp_nys_sum <- hosp_nys %>% 
  group_by(year_week) %>% 
  summarize(across(where(is.numeric), sum, na.rm = TRUE), .groups = "drop") %>% 
  mutate(
    bed_free = bed_capacity - bed_occupied,
    icu_free = icu_capacity - icu_occupied,
    pct_covid = covid_patients / all_patients,
    pct_bed = bed_occupied / bed_capacity,
    pct_icu = icu_occupied / icu_capacity
    )
```

<span class="recent-update">Most recent data: `r month_day_year(hosp_wch_sum$year_week, abbr = TRUE)` -- `r month_day_year(hosp_wch_sum$year_week + days(6), abbr = TRUE)`</span>

In the week ending `r month_day_year(hosp_wch_sum$year_week + days(6), abbr = TRUE)`, there were `r comma(hosp_wch_sum$covid_patients, 1)` patients hospitalized with confirmed or suspected cases of Covid-19 in Westchester hospitals (`r percent(hosp_wch_sum$pct_covid, 1)` of all inpatients). `r percent(hosp_wch_sum$pct_bed, 1)` of all inpatient beds were occupied and `r percent(hosp_wch_sum$pct_icu, 1)` of all ICU beds were occupied. There were `r comma(hosp_wch_sum$bed_free, 1)` total beds free and `r comma(hosp_wch_sum$icu_free, 1)` ICU beds free. Statewide, `r percent(hosp_nys_sum$pct_bed, 1)` of all inpatient beds were occupied and `r percent(hosp_nys_sum$pct_icu, 1)` of all ICU beds were occupied.

<br>

<div class = "l-body-outset">
::::: {.panelset}
::: {.panel}
[Covid-19 patients]{.panel-name}

```{r, layout="body-outset"}
to_plot <- hosp_cap %>% 
  filter(county == "Westchester") %>% 
  group_by(year_week) %>% 
  summarize(across(where(is.numeric), sum, na.rm = TRUE)) %>% 
  mutate(
    bed_pct = bed_occupied / bed_capacity,
    icu_pct = icu_occupied / icu_capacity,
    covid_pct = covid_patients / all_patients
  )

p <- to_plot %>% 
  ggplot(aes(year_week + 3.5, covid_patients)) +
  geom_line(color = "steelblue") +
  geom_point(
    aes(
      text = paste0(
        "<b>", month_day_year(year_week, abbr = TRUE), " - ", month_day_year(year_week + days(6), abbr = TRUE), "</b><br>",
        "Covid-19 patients: ", comma(covid_patients, 1)
        )
      ),
    color = "steelblue"
    ) +
  expand_limits(y = 0) +
  scale_x_date(labels = label_date_short()) +
  scale_y_continuous(labels = label_number_si()) +
  labs(
    title = "Covid-19 patients hospitalized in Westchester",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
  )

ggplotly_config(p) %>% 
  layout(
    hovermode = "x unified",
    annotations = list(
      x = 1, y = -0.175,
      text = 'Source: <a href="https://westchester-covid.mattherman.info/about#covid-19-reported-patient-impact-and-hospital-capacity-by-facility" target="_self">US HHS</a>', 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
:::

::: {.panel}
[% Covid-19 patients]{.panel-name}
```{r, layout="body-outset"}
p <- to_plot %>% 
  ggplot(aes(year_week + 3.5, covid_pct)) +
  geom_line(color = "steelblue") +
  geom_point(
    aes(
      text = paste0(
        "<b>", month_day_year(year_week, abbr = TRUE), " - ", month_day_year(year_week + days(6), abbr = TRUE), "</b><br>",
        "Covid-19 patients: ", percent(covid_pct, 1)
        )
      ),
    color = "steelblue"
    ) +
  expand_limits(y = 0) +
  scale_x_date(labels = label_date_short()) +
  scale_y_continuous(labels = label_percent(1)) +
  labs(
    title = "Covid-19 patients as a percentage of all patients",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
  )

ggplotly_config(p) %>% 
  layout(
    hovermode = "x unified",
    annotations = list(
      x = 1, y = -0.175,
      text = 'Source: <a href="https://westchester-covid.mattherman.info/about#covid-19-reported-patient-impact-and-hospital-capacity-by-facility" target="_self">US HHS</a>', 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
:::

::: {.panel}
[Bed occupancy]{.panel-name}
```{r, layout="body-outset"}
p <- to_plot %>% 
  ggplot(aes(year_week + 3.5, bed_pct)) +
  geom_line(color = "steelblue") +
  geom_point(
    aes(
      text = paste0(
        "<b>", month_day_year(year_week, abbr = TRUE), " - ", month_day_year(year_week + days(6), abbr = TRUE), "</b><br>",
        "Bed occupancy: ", percent(bed_pct, 1), "<br>",
        "Beds occupied: ", comma(bed_occupied, 1), "<br>",
        "Bed capacity: ", comma(bed_capacity, 1), "<br>"
        )
      ),
    color = "steelblue"
    ) +
  expand_limits(y = c(0.5, 1)) +
  scale_x_date(labels = label_date_short()) +
  scale_y_continuous(labels = label_percent(1)) +
  labs(
    title = "Total inpatient bed occupancy at Westchester hospitals",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
  )

ggplotly_config(p) %>% 
  layout(
    hovermode = "x unified",
    annotations = list(
      x = 1, y = -0.175,
      text = 'Source: <a href="https://westchester-covid.mattherman.info/about#covid-19-reported-patient-impact-and-hospital-capacity-by-facility" target="_self">US HHS</a>', 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
:::

::: {.panel}
[ICU bed occupancy]{.panel-name}
```{r, layout="body-outset"}
p <- to_plot %>% 
  ggplot(aes(year_week + 3.5, icu_pct)) +
  geom_line(color = "steelblue") +
  geom_point(
    aes(
      text = paste0(
        "<b>", month_day_year(year_week, abbr = TRUE), " - ", month_day_year(year_week + days(6), abbr = TRUE), "</b><br>",
        "ICU bed occupancy: ", percent(icu_pct, 1), "<br>",
        "ICU beds occupied: ", comma(icu_occupied, 1), "<br>",
        "ICU bed capacity: ", comma(icu_capacity, 1), "<br>"
        )
      ),
    color = "steelblue"
    ) +
  expand_limits(y = c(0, 1)) +
  scale_x_date(labels = label_date_short()) +
  scale_y_continuous(labels = label_percent(1)) +
  labs(
    title = "ICU bed occupancy at Westchester hospitals",
    x = NULL,
    y = NULL
  ) +
  theme_minimal() +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
  )

ggplotly_config(p) %>% 
  layout(
    hovermode = "x unified",
    annotations = list(
      x = 1, y = -0.175,
      text = 'Source: <a href="https://westchester-covid.mattherman.info/about#covid-19-reported-patient-impact-and-hospital-capacity-by-facility" target="_self">US HHS</a>', 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
:::

:::::
</div>

<br>

```{r wch-hosp-table, layout="l-page"}
fips <- c("36071", "36079", "36087", "36119", "09001", "34031", "34003", "34037")

hosp_recent <- hosp_cap %>% 
  filter(
    year_week == max(year_week),
    fips_code %in% fips
    )

to_map <- hosp_geo %>%
  st_drop_geometry() %>% 
  inner_join(hosp_recent, by = "hospital_pk") %>%
  mutate(
    bed_free = bed_capacity - bed_occupied,
    icu_free = icu_capacity - icu_occupied,
    pct_covid = covid_patients / all_patients,
    pct_bed = bed_occupied / bed_capacity,
    pct_icu = icu_occupied / icu_capacity,
    label = paste0(
      "<b>", hospital_name, "</b></br>",
      "Week ending ", month_day_year(year_week + days(6), abbr = TRUE), "<br><br>",
      "Bed occupancy: ", percent(bed_occupied / bed_capacity, 1), "<br>",
      "Beds free: ", comma(bed_free, 1), "</br>",   
      "Total beds: ", comma(bed_capacity, 1), "</br>",
      "</br>",
      "ICU occupancy: ", percent(icu_occupied / icu_capacity, 1), "<br>",
      "ICU beds free: ", comma(icu_free, 1), "</br>",
      "Total ICU beds: ", comma(icu_capacity, 1), "</br>",
      "</br>",
      "Covid-19 patients: ", comma(covid_patients, 1)
      ),
    label = str_replace_all(label, "NA", "&ndash;")
    )

na_format <- c(
    "function(row, data){",
    "  for(var i=0; i<data.length; i++){",
    "    if(data[i+1] === null){",
    "      $('td:eq('+i+')', row).html('NA')",
    "        .css({'color': 'rgb(151,151,151)', 'font-style': 'italic'});",
    "    }",
    "  }",
    "}"  
  )

to_map %>% 
  filter(county == "Westchester") %>% 
  arrange(desc(bed_occupied)) %>% 
  select(
    `Hospital` = hospital_name,
    `City` = city,
    `Covid-19 patients` = covid_patients,
    `% Covid-19 patients` = pct_covid,
    `Beds used` = pct_bed,
    `Total beds` = bed_capacity,
    `ICU beds used` = pct_icu,
    `Total ICU beds` = icu_capacity
    ) %>%
  datatable(
    class = "compact hover",
    extensions = "Responsive",
    caption = tags$caption(
      style = 'caption-side: bottom; text-align: right;',
      withTags(div(HTML('Source: <a href="https://westchester-covid.mattherman.info/about#covid-19-reported-patient-impact-and-hospital-capacity-by-facility" target="_self">US HHS</a>')))
    ),    
    options = list(
      dom = "t",
      rowCallback = na_format,
      columnDefs = list(
        list(visible = FALSE, targets = 0),
        list(className = "dt-head-left", targets = 1:2)
        )
      )
    ) %>% 
  formatRound(c(3, 6, 8), digits = 0) %>% 
  formatPercentage(c(4, 5, 7), digits = 0) %>% 
  formatStyle(1:2, fontFamily = "IBM Plex Sans") %>% 
  formatStyle(3:8, fontFamily = "Roboto Mono")
```

### Regional hospital capacity

```{r hosp_cap_map, out.height="600px", out.width="600px"}
leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addMinicharts(
    layerId = as.character(1:nrow(to_map)),
    lng = to_map$long,
    lat = to_map$lat,
    type = "pie",
    chartdata = select(to_map, `Beds occupied` = bed_occupied, `Beds free` = bed_free),
    width = sqrt(to_map$bed_capacity) * 1.2,
    colorPalette = c("steelblue", "coral"),
    opacity = 0.9,
    popup = popupArgs(html = to_map$label)
    ) %>% 
  addResetMapButton()
```
