---
title: "Westchester Covid-19 Tracking"
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(plotly)
library(sf)
library(leaflet)
library(DT)
library(here)
library(slider)
library(scales)
library(lubridate)
library(leaflet)
library(leaflet.extras)

source(here("R/utils.R"))

```

```{r read_in_data}
nyt_cases <- read_csv(here("data/by-county-cases-deaths-nyt.csv")) %>% 
  group_by(county, state) %>% 
  mutate(
    across(where(is.numeric), 
    list(
      "avg_7" = ~ slide_dbl(.x, mean, .before = 6, .complete = TRUE),
      "avg_14" = ~ slide_dbl(.x, mean, .before = 13, .complete = TRUE)
      )
    )) %>% 
  ungroup()

mun_cases <- read_csv(here("data/by-mun-cases.csv"))
mun_bound <- read_rds(here("data/mun-boundary.rds"))
```

```{r}
recent <- nyt_cases %>% 
  filter(date == max(date), county == "Westchester")

last_week <- nyt_cases %>% 
  filter(date == max(date) - days(7), county == "Westchester")

cases_recent <- recent$new_cases_avg_7 * 7
cases_last_week <- last_week$new_cases_avg_7 * 7

cases_change <- cases_recent - cases_last_week
cases_change_pct <- cases_change / cases_last_week
cases_change_word <- if_else(cases_change > 0, "increase", "decrease")

cases_change_pct_print <- if_else(
  abs(cases_change_pct) < 0.005,
  "small",
  percent(abs(cases_change_pct), 1)
  )

deaths_recent <- recent$new_deaths_avg_7 * 7
deaths_last_week <- last_week$new_deaths_avg_7 * 7

deaths_change <- deaths_recent - deaths_last_week
deaths_change_pct <- deaths_change / deaths_last_week

deaths_change_word <- if_else(deaths_change > 0, "more", "fewer")
```

```{r county_case_bar, layout="l-page"}
to_plot <- nyt_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, new_cases)) +
  geom_col(
    aes(
      text = paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "New cases: ", comma(new_cases, 1), "<br>",
        "7-day average: ", comma(new_cases_avg_7, 1)
        )
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = new_cases_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(label = label_date_short(), date_breaks = "1 month") +
  scale_y_continuous(label = label_number_si()) +
  labs(
    title = "New Cases Reported",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(hovermode = "x unified")
```

<div style = "font-size:1.7rem;line-height:1.2em">
In the last seven days, <span style = "color:white;background-color:steelblue;font-weight:500;font-size:2.2rem;">`r comma(cases_recent, 1)` new cases</span> of Covid-19 have been reported in Westchester County. This is a `r cases_change_pct_print` `r cases_change_word` from the prior seven days.

There were <span style = "color:white;background-color:coral;font-weight:500;font-size:2.2rem;">`r comma(deaths_recent, 1)` new deaths</span> due to Covid-19 reported in the last seven days. This is `r abs(deaths_change)` `r deaths_change_word` than in the prior seven days.
</div>

<div class="index-row l-page">
  <div class="index-column-1">
```{r mun_per_capita_map, out.height="500px", out.width="400px"}
mun_recent <- mun_cases %>% 
  filter(date == max(date), municipality != "Totals") %>%
  mutate(new_cases_day_avg = active_cases / 14)

mun_cases_poly <- mun_bound %>% 
  left_join(mun_recent, by = "municipality") %>% 
  mutate(
    new_cases_per_cap = new_cases_day_avg / total_pop * 1e5,
    total_cases_per_cap = total_cases / total_pop
    )

mun_cases_centroid <- mun_cases_poly %>% 
  st_centroid()

to_map <- mun_cases_poly %>% 
  st_transform(4326) %>% 
  mutate(
    label = paste0(
      "<b>", municipality, "</b><br>",
      "Avg. daily cases: ", comma(new_cases_day_avg, 1), "<br>",
      "Per 100k: ", number(new_cases_per_cap, 0.1)
      )
    )

cd_pal <- colorNumeric(
  palette = "YlOrRd",
  domain = 0:100
  )

to_map %>%
  leaflet(options = leafletOptions(zoomControl = FALSE, minZoom = 10, maxZoom = 10, dragging = FALSE)) %>%
  addPolygons(
    weight = 1.5,
    color = "white",
    fillColor = ~cd_pal(new_cases_per_cap),
    fillOpacity = 0.8,
    label = lapply(to_map$label, htmltools::HTML),
    highlight = highlightOptions(
      weight = 2,
      color = "#666666",
      opacity = 0.7,
      fillOpacity = 0.7,
      bringToFront = TRUE
      )
    ) %>%
  setView(lat = 41.12, lng = -73.75, zoom = 10)
```
  </div>
  
  <div class="index-column-2">
</br>
</br>
</br>
<span style = "font-size:1.7rem;">
[County and metro area &#10230;](county.html)
</span></br>
Westchester county cases, deaths, and tests + comparisons with the metro area

<span style = "font-size:1.7rem;">
[Westchester towns &#10230;](municipality.html)
</span></br>
Recent case rates in the 43 municipalities of Westchester County

<span style = "font-size:1.7rem;">
[Hospital capacity &#10230;](hospital.html)
</span>
</br>
Available beds and Covid-19 patients in metro area hospitals

<span style = "font-size:1.7rem;">
[About the site &#10230;](about.html)
</span></br>
Data sources, etc.

  </div>
</div>

