---
title: "Home"
description: "A tracking site with data about the Covid-19 pandemic in Westchester County, NY. Includes cases by municipality as well as Covid-19 data for hospitals, schools, and nursing homes."
site: distill::distill_website
preview: img/wch-map.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r calc-key-metrics}
recent <- nys_cases %>% 
  filter(date == max(date), county == "Westchester")

last_week <- nys_cases %>% 
  filter(date == max(date) - days(7), county == "Westchester")

recent_nyt <- nyt_cases %>% 
  filter(date == max(date), county == "Westchester")

last_week_nyt <- nyt_cases %>% 
  filter(date == max(date) - days(7), county == "Westchester")

cases_recent <- recent$new_cases_avg_7 * 7
cases_last_week <- last_week$new_cases_avg_7 * 7
cases_change <- cases_recent - cases_last_week
cases_change_pct <- cases_change / cases_last_week
cases_total <- recent$total_cases
cases_change_word <- if_else(cases_change > 0, "increase", "decrease")

cases_change_pct_print <- if_else(
  abs(cases_change_pct) < 0.005,
  "small",
  percent(abs(cases_change_pct), 1)
  )

if (cases_change_pct_print == "small") {
  pct_change_prefix <- "a"
} else {
  pct_change_prefix <- a_or_an(signif(cases_change_pct * 100, 2))
}

deaths_recent <- recent_nyt$new_deaths_avg_7 * 7
deaths_last_week <- last_week_nyt$new_deaths_avg_7 * 7
deaths_change <- deaths_recent - deaths_last_week
deaths_change_pct <- deaths_change / deaths_last_week
deaths_total <- recent_nyt$total_deaths
deaths_change_text <- case_when(
  deaths_change == 0 ~ "the same number</span> as",
  deaths_change > 0 ~ paste(comma(deaths_change, 1), "more</span> than"),
  deaths_change < 0 ~ paste(comma(abs(deaths_change), 1), "fewer</span> than")
  )

tests_recent <- recent$new_tests_avg_7 * 7
tests_last_week <- last_week$new_tests_avg_7 * 7
tests_change <- tests_recent - tests_last_week
tests_change_pct <- tests_change / tests_last_week
tests_total <- recent$total_tests

pos_rate_recent <- cases_recent / tests_recent
pos_rate_last_week <- cases_last_week / tests_last_week
pos_rate_change <- (pos_rate_recent - pos_rate_last_week) * 100

cases_per_cap <- pretty_frac(recent$total_cases / recent_nyt$total_pop)
death_per_cap <- pretty_frac(recent_nyt$total_deaths / recent_nyt$total_pop)
```

<br>
<br>

```{r county_case_bar, layout="l-body-outset"}
to_plot <- nys_cases %>% 
  filter(county == "Westchester") %>% 
  ggplot(aes(date, new_cases)) +
  geom_col(
    aes(
      text = str_replace(paste0(
        "<b>", month_day_year(date, abbr = TRUE), "</b><br>",
        "New cases: ", comma(new_cases, 1), "<br>",
        "7-day average: ", comma(new_cases_avg_7, 1)
        ), "NA", " -")
      ),
    fill = "#82bae8"
    ) +
  geom_line(aes(y = new_cases_avg_7), color = "#3766b3", size = 0.65) +
  scale_x_date(
    label = label_date_short(),
    date_breaks = "2 months",
    limits = c(as.Date("2020-03-01"), NA),
    expand = expansion(add = c(0, 2))
    ) +
  scale_y_continuous(label = label_comma(1)) +
  labs(
    title = "New cases reported",
    x = NULL,
    y = NULL
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major.x = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly_config(to_plot) %>% 
  layout(hovermode = "x")
```

<div class="index-row l-page">
  <div class = "index-column-l l-page">
  
<table class = "key-metric-t" style="width=300;">
  <tr>
    <th class = "t-left"></th>
    <th class = "t-right">Last 7 days</th>
    <th class = "t-right">7-day change</th>
    <th class = "t-right">Total</th>
  </tr>
  <tr>
    <td class = "t-left">Deaths</td>
    <td class = "t-right">`r comma(deaths_recent, 1)`</td>
    <td class = "t-right">`r percent_sign(deaths_change_pct, 1)`</td>
    <td class = "t-right">`r comma(deaths_total, 1)`</td>
  </tr>
  <tr>
    <td class = "t-left">Cases</td>
    <td class = "t-right">`r comma(cases_recent, 1)`</td>
    <td class = "t-right">`r percent_sign(cases_change_pct, 1)`</td>
    <td class = "t-right">`r comma(cases_total, 1)`</td>
  </tr>
  <tr>
    <td class = "t-left">Tests</td>
    <td class = "t-right">`r comma(tests_recent, 1)`</td>
    <td class = "t-right">`r percent_sign(tests_change_pct, 1)`</td>
    <td class = "t-right">`r label_number_si(0.01)(tests_total)`</td>
    
  </tr>
    <tr>
    <td class = "t-left">Pos rate</td>
    <td class = "t-right">`r percent(pos_rate_recent, 0.1)`</td>
    <td class = "t-right">`r num_sign(pos_rate_change, 0.1)`</td>
    <td class = "t-right"></td>
  </tr>
  
</table>
  </div>

  <div class = "index-column-r l-page" style = "font-size:1.2rem;line-height:1.2em;font-family:'IBM Plex Sans', sans-serif;">

In the last seven days, <span style = "color:white;background-color:steelblue;">`r comma(cases_recent, 1)` new cases</span> of Covid-19 have been reported in Westchester County. This is `r pct_change_prefix` <span style = "color:white;background-color:steelblue;">`r cases_change_pct_print` `r cases_change_word`</span> from the prior seven days. Since the beginning of the pandemic, <span style = "color:white;background-color:steelblue;">`r cases_per_cap`</span> Westchester residents have tested positive for the novel coronavirus.

There were <span style = "color:white;background-color:coral;">`r comma(deaths_recent, 1)` new deaths</span> due to Covid-19 reported in the last seven days. This is <span style = "color:white;background-color:coral;">`r deaths_change_text` in the prior seven days. <span style = "color:white;background-color:coral;">`r death_per_cap`</span> Westchester residents have died from Covid-19.
  </div>
</div>

<br>

<div class="index-row l-page">
  <div class="index-column-l l-page">
```{r mun_per_capita_map, out.height="500px", out.width="400px"}
mun_recent <- mun_cases %>%
  filter(date == max(date), municipality != "Totals") %>%
  mutate(new_cases_day_avg = active_cases / 14)

mun_cases_poly <- mun_bound %>%
  left_join(mun_recent, by = "municipality") %>%
  mutate(
    new_cases_per_cap = new_cases_day_avg / total_pop * 1e5,
    total_cases_per_cap = total_cases / total_pop
    )

to_map <- mun_cases_poly %>% 
  st_transform(4326) %>% 
  mutate(
    label = paste0(
      "<b>", municipality, "</b><br>",
      "Avg. daily cases: ", comma(new_cases_day_avg, 0.1), "<br>",
      "Per 100k: ", number(new_cases_per_cap, 0.1)
      ),
    new_cases_per_cap_map = if_else(new_cases_per_cap > 125, 125, new_cases_per_cap)
    )

cd_pal <- colorNumeric(
  palette = "YlOrRd",
  domain = 0:125
  )

to_map %>%
  leaflet(
    options = leafletOptions(
      zoomControl = FALSE,
      minZoom = 10,
      maxZoom = 10,
      dragging = FALSE
      )
    ) %>%
  addPolygons(
    weight = 1.5,
    color = "white",
    fillColor = ~cd_pal(new_cases_per_cap_map),
    fillOpacity = 0.8,
    label = lapply(to_map$label, htmltools::HTML),
    highlight = highlightOptions(
      weight = 2,
      color = "#666666",
      opacity = 0.7,
      fillOpacity = 0.7,
      bringToFront = TRUE
      )
    ) %>%
  setView(lat = 41.12, lng = -73.75, zoom = 10)
```
  </div>

  <div class="index-column-r l-page">
</br>
</br>
</br>
    <div style = "font-family:'IBM Plex Sans', sans-serif;">
<span style = "font-size:1.6rem;">
[County and metro area &#10230;](county.html)
</span></br>
Cases, deaths, and tests; cases in schools; racial disparities; nursing home deaths

<span style = "font-size:1.6rem;">
[Westchester municipalities &#10230;](municipality.html)
</span></br>
Recent cases in the 43 municipalities of Westchester County

<span style = "font-size:1.6rem;">
[Hospital capacity &#10230;](hospital.html)
</span>
</br>
Available beds and Covid-19 patients in metro area hospitals

<span style = "font-size:1.6rem;">
[About the site &#10230;](about.html)
</span></br>
Data sources, etc.
    </div>
  </div>
</div>

<br>
<br>

<p style="text-align: center;">*Updated `r month_day_year(Sys.Date(), abbr = TRUE)`, `r pretty_time(Sys.time())`*</p>

```{css}
d-title, d-byline {
  display: none
}
```

