---
title: "Westchester municipalities"
site: distill::distill_website
preview: img/wch-map.png
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

```{r}
mun_recent <- mun_cases %>% 
  filter(date == max(date), municipality != "Totals") %>%
  mutate(new_cases_day_avg = active_cases / 14)

mun_cases_poly <- mun_bound %>% 
  left_join(mun_recent, by = "municipality") %>% 
  mutate(
    new_cases_per_cap = new_cases_day_avg / total_pop * 1e5,
    total_cases_per_cap = total_cases / total_pop
    ) %>% 
  arrange(desc(new_cases_per_cap))

tot_cases_least <- mun_cases_poly %>% 
  arrange(total_cases_per_cap)

tot_cases_most <- mun_cases_poly %>% 
  arrange(desc(total_cases_per_cap))
```

<span class="recent-update">Most recent data: `r month_day_year(max(mun_recent$date), abbr = TRUE)`</span>

The municipality with the most new cases per capita in the last 14 days is **`r mun_cases_poly$municipality[1]`**, where there have been an average of `r comma(mun_cases_poly$active_cases[1] / 14, 1)` new cases per day reported between `r month_day_year(max(mun_recent$date) - days(13), abbr = TRUE)` and `r month_day_year(max(mun_recent$date), abbr = TRUE)`  (`r comma(mun_cases_poly$new_cases_per_cap[1], 1)` per 100,000 residents). **`r mun_cases_poly$municipality[2]`** is another recent hot spot, with `r comma(mun_cases_poly$active_cases[2] / 14, 1)` new cases per day reported in the last two weeks (`r comma(mun_cases_poly$new_cases_per_cap[2], 1)` per 100,000). 

Throughout the pandemic, the municipalities that have had the largest share of their population test positive for coronavirus are **`r tot_cases_most$municipality[1]`** (`r pretty_frac(tot_cases_most$total_cases_per_cap[1])`) and **`r tot_cases_most$municipality[2]`** (`r pretty_frac(tot_cases_most$total_cases_per_cap[2])`). The municipalities that have had the fewest cases per capita are **`r tot_cases_least$municipality[1]`** (`r pretty_frac(tot_cases_least$total_cases_per_cap[1])`) and **`r tot_cases_least$municipality[2]`** (`r pretty_frac(tot_cases_least$total_cases_per_cap[2])`).

### New cases per 100,000 residents (14-day average)

```{r mun_per_capita_map, out.height="600px", out.width="600px"}
mun_cases_centroid <- mun_cases_poly %>% 
  st_centroid()

to_map <- mun_cases_poly %>% 
  st_transform(4326) %>% 
  mutate(
    label = paste0(
      "<b>", municipality, "</b><br>",
      "Avg. daily cases: ", comma(new_cases_day_avg, 0.1), "<br>",
      "Per 100k: ", number(new_cases_per_cap, 0.1)
      ),
    new_cases_per_cap_map = if_else(new_cases_per_cap > 125, 125, new_cases_per_cap)
    )

to_map_total <- mun_cases_centroid %>% 
  st_transform(4326) %>% 
  mutate(
    label = paste0(
      "<b>", municipality, "</b><br>",
      "Total cases: ", comma(total_cases, 1), "<br>",
      "Share of population: ", pretty_frac(total_cases_per_cap)
      )
    )

cd_pal <- colorNumeric(
  palette = "YlOrRd",
  domain = 0:125
  )

leaflet(to_map) %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addPolygons(
    group = "Cases per 100k",
    weight = 1.5,
    color = "white",
    fillColor = ~cd_pal(new_cases_per_cap_map),
    fillOpacity = 0.8,
    label = lapply(to_map$label, htmltools::HTML),
    highlight = highlightOptions(
      weight = 2,
      color = "#666666",
      opacity = 0.7,
      fillOpacity = 0.7,
      bringToFront = TRUE
      )
    ) %>%
  addCircles(
    data = to_map_total,
    group = "Total cases",
    radius = ~ sqrt(total_cases) * 50,
    color = "steelblue",
    opacity = 0.7,
    weight = 1.5,
    fillColor = "steelblue",
    fillOpacity = 0.2,
    label = lapply(to_map_total$label, htmltools::HTML),
    highlightOptions = highlightOptions(sendToBack = TRUE, color = "#6fb6f2", weight = 2.5)    
    ) %>% 
  addLegend(
    group = "Cases per 100k",
    pal = cd_pal,
    values = 0:125,
    opacity = 0.7,
    bins = c(0, 25, 50, 75, 100, 125),
    position = "bottomright",
    title = "Avg. daily cases per 100k<br>in last 14 days"
    ) %>% 
  addLayersControl(
    baseGroups = c("Cases per 100k", "Total cases"),
    options = layersControlOptions(collapsed = FALSE)
   ) %>% 
  setView(lat = 41.1, lng = -73.75, zoom = 10) %>% 
  addResetMapButton()
```

<br>

```{r town-table, layout="l-body-outset", out.height="510px"}
mun_cases_poly %>% 
  st_drop_geometry() %>% 
  arrange(desc(new_cases_per_cap)) %>% 
  mutate(per_cap_frac = pretty_frac(total_cases_per_cap)) %>% 
  select(
    `Municipality` = municipality,
    `Avg daily cases per 100k` = new_cases_per_cap,
    `Avg daily cases` = new_cases_day_avg,
    `Total cases` = total_cases,
    `Share of pop` = per_cap_frac,
    `Total pop` = total_pop,
    total_cases_per_cap
    ) %>% 
  datatable(
    class = "compact hover",
    extensions = c("Scroller", "Responsive"),
    caption = tags$caption(
      style = 'caption-side: bottom; text-align: right;',
      withTags(div(HTML('Source: <a href="https://westchester-covid.mattherman.info/about#westchester-county-covid-19-dashboard" target="_self">Westchester County DOH</a>')))
    ),
    options = list(
      dom = "ft",
      deferRender = TRUE,
      scrollY = 250,
      scroller = TRUE,
      columnDefs = list(
        list(orderData = 7, targets = 5),
        list(visible = FALSE, targets = c(0, 7)),
        list(className = "dt-head-left", targets = 1),
        list(className = "dt-center", targets = 5)
        )
      )
    ) %>% 
  formatRound(2:3, digits = 1) %>% 
  formatRound(c(4, 6), digits = 0) %>% 
  formatStyle(1, fontFamily = "IBM Plex Sans") %>% 
  formatStyle(2:6, fontFamily = "Roboto Mono")
```

<br>

## Municipal trends

Westchester County does not currently publish historical data at the municipal-level, but I have reconstructed this data using maps [posted on the WestchesterGov Twitter account](https://twitter.com/westchestergov/status/1336045976981811206). The earliest maps that included the number of active cases by municipality were posted in early May. As such, the data I have *largely misses the first wave of the pandemic* from March and April 2020.

<aside>
Because of the unavailability of this historical data from the County, caution should be used when interpreting municipal-level trends.
</aside>

The chart below shows the severity of the pandemic in each municipality by plotting the rolling 14-day average of new cases per 100,000 residents by day. The municipalities are ordered from top to bottom by their current average daily cases per 100,000.

<aside>
For days with missing data, I have imputed the case numbers using linear interpolation.
</aside>

<div class="l-page heat-desktop"> 
```{r mun_heat_map, layout="l-page", out.height="700px"}
all_dates <- seq(min(mun_cases$date), max(mun_cases$date), by = "1 day")
all_mun <- mun_bound$municipality

all_days_mun <- crossing(date = all_dates, municipality = all_mun) %>% 
  filter(date >= as.Date("2020-05-05")) %>% 
  left_join(mun_cases, by = c("date", "municipality")) %>% 
  left_join(st_drop_geometry(mun_bound), by = "municipality") %>% 
  mutate(active_per_cap = active_cases / total_pop) %>% 
  group_by(municipality) %>% 
  mutate(
    active_cases_est = na.approx(active_cases),
    active_per_cap_est = active_cases_est / total_pop,
    active_per_cap_limit = if_else(active_per_cap_est / 14 * 1e5 > 125, 125, active_per_cap_est / 14 * 1e5)
    ) %>% 
  ungroup()

p <- all_days_mun %>%  
  mutate(municipality = fct_reorder(municipality, active_per_cap_est, last)) %>% 
  ggplot(aes(as.numeric(as.POSIXct(date)) * 1000, municipality, fill = active_per_cap_limit)) + #https://stackoverflow.com/questions/38919395/plotly-r-cant-make-custom-xaxis-date-ranges-to-works
  geom_tile(
    aes(
      text = str_replace_all(paste0(
        "<b>", municipality, " - ", month_day_year(date, abbr = TRUE), "</b><br>",
        "Avg. daily cases: ", comma(active_cases_est / 14, 0.1), "<br>",
        "Per 100k: ", comma(active_per_cap_est / 14 * 1e5, 0.1)
        ), "NA", "missing")
      )
    ) +
  scale_fill_distiller(
    palette = "YlOrRd",
    direction = 1
    ) +
  scale_x_continuous(expand = expansion(mult = c(0.01, 0))) +
  labs(x = NULL, y = NULL, title = "Average daily cases per 100k (beginning May 2020)") +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", size = 20)
    )

ggplotly(p, tooltip = "text") %>%
  layout(
    xaxis = list(
      type = "date",
      tickformat = "%b %-d"
    ),
    yaxis = list(fixedrange = TRUE),
    font = list(family = "IBM Plex Sans Condensed"),
    hoverlabel = list(font = list(family = "IBM Plex Sans Condensed"), align = "left"),
    annotations = list(
      x = 1, y = -0.06,
      text = 'Source: <a href="https://westchester-covid.mattherman.info/about#westchester-county-covid-19-dashboard" target="_self">Westchester County DOH</a>', 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    ) %>%
    plotly::config(
      displayModeBar = TRUE,
      displaylogo = FALSE,
      modeBarButtonsToRemove = c("zoom2d", "pan2d", "zoomIn2d", "zoomOut2d",
                                 "autoScale2d", "toggleSpikelines", "hoverClosestCartesian",
                                 "hoverCompareCartesian", "toImage")
      )
```
</div>

<div class="l-page heat-mobile"> 
```{r mun_heat_map_mobile, layout="l-page", out.height="700px"}
p <- all_days_mun %>%  
  mutate(municipality = fct_reorder(municipality, active_per_cap_est, last)) %>% 
  ggplot(aes(date, municipality, fill = active_per_cap_est)) +
  geom_tile(
    aes(
      text = str_replace_all(paste0(
        "<b>", municipality, " - ", month_day_year(date, abbr = TRUE), "</b><br>",
        "Avg. daily cases: ", comma(active_cases_est / 14, 0.1), "<br>",
        "Per 100k: ", comma(active_per_cap_est / 14 * 1e5, 0.1)
        ), "NA", "missing")
      )
    ) +
  scale_fill_distiller(
    palette = "YlOrRd",
    direction = 1,
    ) +
  scale_x_date(
    labels = label_date_short(),
    breaks = "1 month",
    expand = expansion(mult = c(0.01, 0)),
    limits = c(as.Date("2020-10-01"), NA)
    ) +
  labs(x = NULL, y = NULL, title = "Avg daily cases per 100k (beginning Oct 2020)") +
  theme_minimal() +
  theme(
    legend.position = "none",
    panel.grid = element_blank(),
    plot.title = element_text(face = "bold", size = 16)
    )

ggplotly_config(p) %>% 
  layout(
    annotations = list(
      x = 1, y = -0.075,
      text = 'Source: <a href="https://westchester-covid.mattherman.info/about#westchester-county-covid-19-dashboard" target="_self">Westchester County DOH</a>', 
      showarrow = FALSE,
      xref = "paper", yref = "paper", 
      xanchor = "right", yanchor = "auto",
      xshift = 0, yshift = 0
      )
    )
```
</div>

<br>

```{r}
mun_acs_calc <- mun_acs %>% 
  left_join(mun_recent, by = "municipality") %>% 
  mutate(total_cases_per_cap = total_cases / total_pop)

cors <- mun_acs_calc %>% 
  select(total_cases_per_cap, poverty_pct, crowded_pct, black_lat_pct) %>% 
  cor() %>%
  as.data.frame()

mun_sd <- highlight_key(mun_acs_calc, ~municipality, group = "Municpality")

# mun_sd <- mun_acs_calc %>% 
#   SharedData$new()

p <- mun_sd %>% 
  ggplot(aes(poverty_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Families below 200% poverty line: ", percent(poverty_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.02) + 
  scale_x_continuous(label = label_percent(1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.2, by = 0.02)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
  labs(
    x = "Families below 200% of poverty line",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p_plotly <- ggplotly_config(p)

p2 <- mun_sd %>%
  ggplot(aes(crowded_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Overcrowded households: ", percent(crowded_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.02) + 
  scale_x_continuous(label = label_percent(1), breaks = c(0, 0.05, 0.1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.2, by = 0.02)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
  labs(
    x = "Overcrowded households",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p2_plotly <- ggplotly_config(p2)

p3 <- mun_sd %>%
  ggplot(aes(black_lat_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Black or Latino population: ", percent(black_lat_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.02) + 
  scale_x_continuous(label = label_percent(1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.2, by = 0.02)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
    labs(
    x = "Black or Latino population",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p3_plotly <- ggplotly_config(p3)

p4 <- mun_sd %>%
  ggplot(aes(hs_less_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Black or Latino population: ", percent(hs_less_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.02) + 
  scale_x_continuous(label = label_percent(1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.2, by = 0.02)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
    labs(
    x = "Black or Latino population",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p4_plotly <- ggplotly_config(p4)

scatter <- list(p_plotly, p2_plotly, p3_plotly)
```

## Socioeconomic factors associated with Covid-19 prevalance

In the same way that the [death rate from Covid-19 is Westchester County is higher for Black and Latino people](county.html#racial-disparities) than for White people, the total number of cases in the community is also [correlated with socioeconomic characteristics](https://www.cdc.gov/coronavirus/2019-ncov/community/health-equity/race-ethnicity.html). The risk of being infected with the novel coronavirus is higher for people of color due to factors including disproportionate representation in "essential" occupations, overcrowded and multigenerational households, and systemic discrimination and racism resulting in disparities in education and income as well as reduced access to high quality healthcare.

The scatterplots below illustrate the relationship between the total cases per capita in each municipality and three selected characteristics: poverty, overcrowding, and Black or Latino population. Each circle represents one of the 43 Westchester municipalities. Circles in the upper right-hand corner of each plot represent municipalities with high case rates and high levels of each socioeconomic variable. Municipalities with low case rates and low levels of each socioeconomic factor are in the bottom left.

<aside>
[See here](about.html#american-community-survey) for more information on how each of the socioeconomic variables is calculated.
<br><br>
The size of each circle is based on the total population of each municipality.
</aside>

There is a strong positive correlation between each of the variables and the number of cases per capita. This suggests that the national trends of disproportionate impacts of the pandemic on disadvantaged communities also exists in Westchester County.

<aside>
Hover over a circle to see which municipality it represents.
</aside>

<div class="scatter-wide l-page">
```{r, layout="l-page", out.height="400px", eval=T}
subplot(scatter, titleX = TRUE, shareY = TRUE) %>%
  highlight(
    on = "plotly_hover",
    off = "plotly_doubleclick",
    opacityDim = 0.25,
    selectize = TRUE
    )
```
</div>

```{r}
p <- mun_acs_calc %>% 
  ggplot(aes(poverty_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Families below 200% poverty line: ", percent(poverty_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.02) + 
  scale_x_continuous(label = label_percent(1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.2, by = 0.02)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
  labs(
    x = "Families below 200% of poverty line",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p_plotly <- ggplotly_config(p)

p2 <- mun_acs_calc %>%
  ggplot(aes(crowded_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Overcrowded households: ", percent(crowded_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.02) + 
  scale_x_continuous(label = label_percent(1), breaks = c(0, 0.05, 0.1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.2, by = 0.02)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
  labs(
    x = "Overcrowded households",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p2_plotly <- ggplotly_config(p2)

p3 <- mun_acs_calc %>%
  ggplot(aes(black_lat_pct, total_cases_per_cap)) +
  geom_point(
    aes(
      size = total_pop,
      text = paste0(
        "<b>", municipality, "</b><br>",
        "Black or Latino population: ", percent(black_lat_pct, 1), "<br>",
        "Share of population with positive test: ", pretty_frac(total_cases_per_cap)
        )
      ),
    shape = 21,
    color = "steelblue",
    fill = adjustcolor("steelblue", alpha.f = 0.3),
    alpha = 0.9
    ) +
  expand_limits(x = 0, y = 0.02) + 
  scale_x_continuous(label = label_percent(1)) +
  scale_y_continuous(label = label_percent(1), breaks = seq(0, 0.2, by = 0.02)) + 
  scale_size(trans = "log10", range = c(1, 5)) +
  labs(
    x = "Black or Latino population",
    y = "Total population with positive test"
  ) +
  theme_minimal() +
  theme(axis.title.x = element_text(size = 14))

p3_plotly <- ggplotly_config(p3)
```

<div class="scatter-long">
```{r eval=T}
p_plotly
```

```{r eval=T}
p2_plotly
```

```{r eval=T}
p3_plotly
```
</div>